---
title: "assignment 4_task 2"
author: "Craig Stuart"
date: "11/13/2020"
output: html_document
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#attach all necessary packages
library(tidyverse)
library(here)
```

#### Read in Bonanza Creek data
```{r}
bonanza_creek <- read_csv(here("data", "bonanza_hares.csv"))
```

## 1. Introduction
- 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report.

## 2. Data and Analyses
- A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used

## 3. Exploratory Findings

### - Annual juvenile Snowshoe hare trap counts

### - Visualize juvenile Snowshoe hare weights 

### - Juvenile weight comparison

### - Relationship between juvenile weight & hind foot length

## 4. Summary
- Include a very short summary describing your main findings from your juvenile hares exploration (bulletpoints are fine)

## 5. Citations
- Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14







