---
title: "assignment 4_task 2: Snowshoe Hare: Exploratory Report"
author: "Craig Stuart"
date: "11/13/2020"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# attach all necessary packages

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(knitr)
library(effsize)
library(ggbeeswarm)
library(kableExtra)
library(broom)
```

```{r}
# Read in Bonanza Creek data

bonanza_creek <- read_csv(here("data", "bonanza_hares.csv"))
```

## 1. Introduction
- This report provides an exploratory overview of differences in weight and hind foot length of male and female juvenile snowshoe hares observed in the Bonanza Creek Experimental Forest, Long Term Ecological Research (LTER) study from 1999 - 2012 (Kielland et. al. (2017). First, the total trap counts of all juvenile snowshoe hares throughout the study are summarized. Next, the weights of male and female snowshoe hares are compared across the three grids (sites) to determine if snowshoe hare weights differ between sex and grid. Finally, it also explores if a relationship exists between juvenile snowshoe hare weight and hind foot length to determine if there is a correlation between the two physical features. Major findings are posted in the results section at the end of the report.  

## 2. Data and Analyses
- Physical data on snowshoe hare were provided by a study conducted by Kielland et. al. (2017) and the Bonanza Creek Long Term Ecological Research group. The study collected capture-recapture data on snowshoe hare's at three(3) grid, or sites, (bonrip, bonbs, bonmat) within the Bonanza Creek Experimental Forest between 1999 and 2012. The data contains 3197 observations over 14 variables, including grid, sex, age, and measurements of weight (grams) and hind foot length (millimeters). We explore the visualizations of juvenile hare trap counts and the comparisons of juvenile hare weights between sex and grid. Following exploratory data visualizations, male and female weight comparisons are analyzed statistically with a two sample t-test using a significance level ($\alpha$) of 0.05. Cohen's d is then used to describe the differences between snowshoe hare sex. Finally, the relationship between juvenile weight and high foot length is explored by a simple linear regression. All analyses are conducted using R version 4.0.2 and RStudio Version 1.3.1073.

## 3. Exploratory Findings

### 3.1: Annual juvenile Snowshoe hare trap counts (Figure 1)

```{r}
# Find the total number of juvenile hare trappings for each year of the study
# `filer()` for juvenile observations, `mutate()` the date to get a year variable, and summarize the `count()` by year

juvenile_trap <- bonanza_creek %>% 
  filter(age == "j") %>% #filter to match only juvenile observations
  mutate(date_new = mdy(date)) %>% #reformat date as mdy
  mutate(year = year(date_new)) %>% #add column with year only
  count(year) #summarize counts per year
```

```{r}
# Use `ggplot()` to visualize the absolute counts per year

ggplot(data = juvenile_trap, aes(x = year, y = n)) +
  geom_bar(stat = "identity", fill = "blue", color = "black") + #select bar graph and add aesthetics
  labs(x = "Study Year", #add x, y, and title labels
       y = "Number of Counts", 
       title = "Annual Trap Counts") +
  scale_y_continuous(breaks = seq(0,130,20)) + #adjust y axis values
  scale_x_continuous(breaks = seq(1999, 2012, 1))+ #adjust x axis values
  theme_light() + #choose theme
  theme(plot.title = element_text(hjust = 0.5)) #center plot title
```

**Figure 1**: Absolute trap counts of juvenile snowshoe hare at all grids for each year of the Bonanza Creek Long Term Ecological Research (LTER) Experimental Forest study (1999-2012). Data: Kielland et. al. (2017).

```{r, include=FALSE}
# Use `summary()` to display mean, median, max, and min

juvenile_trap_summary <- summary(juvenile_trap)
```

- The maximum number of juvenile hare trappings occurred in 1999 (n = 126) and the minimum juvenile hare trappings occurred in 2010 (n = 2). The mean hare trappings throughout the entire study was 31.5 and the median was 18.5. Some general trends to note would be that the number of juvenile hare trappings had high variance from year to year. Also, that it's hard to tell for if there were zero hare trappings in 2002 and 2009 or if there were no study conducted. What this graph fails to present is the amount of effort put into each year. For instance, it would make sense if in 2010 only two days resulted in 2 trap counts, but the data would be very different if only 2 trap counts were observed in 365 days of the study. Therefore, going forward, a better way to standardize the juvenile hare population is to calculate the trap count per unit effort.

### 3.2: Visualize juvenile Snowshoe hare weights 

```{r}
# Wrangle `bonanza_creek` data to include observations of juvenile weight and sex. 
# `select()` desired variables weight, age, and sex, and `filter()` juvenile observations

j_weights_comp <- bonanza_creek %>% 
  select("weight", "age", "sex", "grid") %>% #select variables weight, age, sex, and grid
  filter(age == "j") %>%  #filter only juvenile observations
  mutate(sex = case_when( #replace sex designations with full names, male and female
    sex %in% "f" ~ "Female",
    sex %in% "m" ~ "Male"))

# plot juvenile hare weight by sex and grid on a geom_beeswarm

ggplot(data = j_weights_comp, aes(x = sex, y = weight)) +
  geom_beeswarm(aes(color = sex)) + #select beeswarm plot
  facet_wrap(~grid) + #facet wrap by grid
  geom_boxplot(fill = NA, width = 0.5, outlier.color = NA) + #add box end points around the data percentile
  theme_bw() + #choose theme
  labs(x = "Sex", #add x, y, and title labels
       y = "Weight (g)",
       title = "Juvenile Male and Female Weight Comparison") +
  theme(plot.title = element_text(hjust = 0.5)) #center plot title
```

**Figure 2**: Male and female snowshoe hare weight (g) observations for three grids (bonds, bonmat, bonrip) in the Bonanza Creek LTER Experimental Forest study. Male (teal) and female (red) indicate individual observations for weight (g) of juvenile snowshoe hares. Box endpoints indicate the 25th and 75th percentile values. Data: Kielland et. al. (2017).

- The beeswarm plot helps to visualize the differences of the weight distributions among juvenile snowshoe hare sex for each grid. From this plot, it is difficult to determine if the data is normally distributed and therefore we would want to conduct further tests (histogram, quantile-quantile plot, or `pnorm()`) to determine normalcy of the data. We can generally make out that male juvenile hare weights tend to be greater than females at all three grids. 

### 3.3: Juvenile weight comparison

```{r, include=FALSE}
# Wrangle `j_weights_sex` data to remove any NA weights
# `summarize()` the mean, sd, and sample size of male and female juvenile hare

j_weight_sex <- bonanza_creek %>% 
  select("weight", "sex", "age") %>% #select variables weight, sex, and age
  filter(age == "j") #filter only juvenile observations

# get summary statistics

sex_weight_stat <- j_weight_sex %>% 
  filter(weight != "NA") %>% #remove NA weight variables 
  group_by(sex) %>% #group by male and female
  summarize(
    mean_weight = mean(weight), #calculate mean
    sd_weight = sd(weight), #calculate standard deviation
    sample_size = n()) #calculate sample size

# visualize data distribution with histogram

ggplot(data = j_weight_sex, aes(x = weight)) +
  geom_histogram() +
  facet_wrap(~sex)

# visualize data distribution with quantile-quantile plot

ggplot(data = j_weight_sex, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
```

**Table 1.** Descriptive statistics (mean, standard deviation, and sample size) for juvenile male and female snowshoe hare weights.  
```{r}
# Display summary statistics in a tabular form

sex_weight_stat %>% 
  kable(col.names = c("Sex", "Mean Hare Weight (g)", "Standard deviation (g)", "Sample size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

```{r, include=FALSE}
#Two sample t-test and cohen's d of male and female weights

# Use `pull()` to create vectors of male and female weights
# Run a two sample `t.test()`

m_weight <- j_weight_sex %>% 
  filter(sex == "m") %>% #filter only male observations
  pull(weight) #put male weights into a vector

f_weight <- j_weight_sex %>% 
  filter(weight != "NA") %>% #remove NA weight observations
  filter(sex == "f") %>% #filter only female observations
  pull(weight) #put female weight into a vector

# Two sample t-test

hare_weight_ttest <- t.test(m_weight, f_weight)
hare_weight_ttest

# Check distributions

hist(m_weight)
hist(f_weight)

# Run a cohen's.d test to determine effect size

cohen.d_weight <- cohen.d(m_weight, f_weight)
```

- On average, the mean weight (grams) of juvenile male snowshoe hares is greater than female snowshoe hares (945.9 $\pm$ 333.2 (n = 163) and 855.4 $\pm$ 292.3 (n = 197), respectively). While the absolute difference between means is 90.5 grams (a 10.0% difference), the difference in means is significant (Welch's two-sample t-test: t(2.7096) = 325.02, p = 0.007), and the effect size is small (Cohen's d = 0.29).

### 3.4: Relationship between juvenile weight & hind foot length (Figures 3 and 4)

```{r}
# Explore the relationship between juvenile Snowshoe hare hind foot length and weight

footlength_weight <- bonanza_creek %>% 
  select("age", "weight", "hindft") %>% #select variables age, weight, and hind foot
  filter(age == "j") #filter only juvenile observations

ggplot(data = footlength_weight, aes(x = hindft, y = weight)) +
  geom_point(color = "purple4") + #select point plot
  labs(x = "Hind Foot Length (mm)", #label x, y, and title
       y = "Weight (g)",
       title = "Juvenile Snowshoe Hare Hind Foot Length vs. Weight ")+
  theme_light() + #choose theme
  theme(plot.title = element_text(hjust = 0.5)) #center plot title
```

**Figure 3**: General relationship between juvenile snowshoe hare hind foot length (mm) and weight (g) from Bonanza Creek LTER Experimental Forest study. Data: Kielland et. al. (2017).

```{r, include=FALSE}
# linear regression

weight_hindft_lm <- lm(weight~hindft, data = footlength_weight)
summary(weight_hindft_lm)

# Get tidy versions of the model output to call later on in text

footlength_weight_lm_tidy <- tidy(weight_hindft_lm)
footlength_weight_lm_glance <- glance(weight_hindft_lm)

# Plot linear regression

plot(weight_hindft_lm)

# Get Pearson's r correlation

snowshoe_cor <- cor.test(footlength_weight$weight, footlength_weight$hindft)

# Tidy version of correlation output:

snowshoe_cor_tidy <- tidy(snowshoe_cor)
snowshoe_cor_tidy
```

```{r, include=FALSE}
# Plot linear regression model

ggplot(data = footlength_weight, aes(x = hindft, y = weight)) +
  geom_point(color = "purple4") +
  labs(x = "Hind Foot Length (mm)", y = "Weight (g)", title = "Juvenile Hare Hind Foot Length vs Weight") +
  geom_smooth(method = "lm", se = FALSE, color = "gray30") +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5))
```

- Figure 3 shows the general relationship between hind food length and weight of juvenile snowshoe hares. The relationship appears to be slightly linear but multiple outliers appear as hind foot length increases. A simple linear regression reveals a somewhat linear relationship (p < 0.001, R^2^ = 0.30) with an average slope of $\beta$ = 9.52 g mm^-1^ (i.e., for each one millimeter increase in hind foot length we expect an average increase in hare weight of 9.52 g). The relationship between hind foot length and weight of juvenile Snowshoe hare is moderately, and significantly, positively correlated (Pearson's *r* = 0.55, p < 0.001). However, diagnostic plots (not included in report) reveal that the assumption of homoscedasticity is violated because the data have a heteroscedastic spread of residuals about the model. Even though the linear regression statistics point to a significant and positive linear relationship, the assumptions tell us otherwise, Therefore, I chose not to include the plot containing the linear regression because I did not feel as this was the most responsible way to describe the relationship. 

## 4. Summary

Exploratory data analysis reveals the following:

- Male snowshoe hare mean weight is greater than female snowshoe hare mean weight at all three grids in the study.
- Only 30% of variance in snowshoe hare weight is described by hind foot length.
- Although a slight linear relationship with a slope (expected average increase in weight with a 1 mm incense in hind foot length) of 9.52 g mm^-1^ is found, the violation of the homoscedasticity assumption  prevents us from using hind foot length as a variable to explain a change in weight. 

Next steps:

- Calculate trap per unit effort for annual hare trap counts.
- Compare hind foot length and weight relationship between sex and grids. 

## 5. Citations
- Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14







